---
title: '405'
author: "Xintong Cai"
date: "2025-04-13"
output:
  pdf_document: default
  html_document: default
---
## 1.Assessing estimates of the 95th percentile of 100 iid normal(0,1) random variables.

### (a)
```{r}
simple_95th <- function(x) {
  if (length(x) != 100) {
    stop("Input vector must be of length 100.")
  }
  
  # Sort the vector
  sorted_x <- sort(x)
  
  # Return the 95th value
  return(sorted_x[95])
}
```


### (b)

```{r}
simple_96th <- function(x) {
  if (length(x) != 100) {
    stop("Input vector must be of length 100.")
  }
  
  # Sort the vector
  sorted_x <- sort(x)
  
  # Return the 95th value
  return(sorted_x[96])
}
```

### (c)

```{r}
simple_avg_9596 <- function(x) {
  if (length(x) != 100) {
    stop("Input vector must be of length 100.")
  }
  
  # Sort the vector
  sorted_x <- sort(x)
  
  # Return the 95th value
  return(mean(sorted_x[95:96]))
}
```

### (d)For	each	of	your	functions	in	parts	a-c,	as well	as	the	function quantile(x,0.95),	do	the	following:
For (a) function 
(i)
```{r}
x <- rnorm(100, mean = 0, sd = 1)

quantile(x, 0.95)
```


```{r}
simple_95th(x)
```
(ii) Repeat the steps for 100,000 times
```{r}
n_sim <- 100000
simple_95th_vals <- numeric(n_sim)
quantile_95th_vals <- numeric(n_sim)
sample_mean <- numeric(n_sim)

# Simulation loop
for (i in 1:n_sim) {
  x <- rnorm(100)
  simple_95th_vals[i] <- simple_95th(x) # the real 95th generation
  quantile_95th_vals[i] <- quantile(x, 0.95)  # default method
  sample_mean[i] <- mean(quantile_95th_vals[1:i])
}

# Summarize the results
cat("Mean of simple 95th estimate:", mean(simple_95th_vals), "\n")
```

(iii)Plot the sample mean of the fist m of my estimates, which is the quantile_95th_vals, and I smooth the plot by taking 100 as the interval to smooth it. The blue line is the interpolation while the red is the true line. 

```{r}
smoothed_index <- seq(100, 100000, by = 100)
smoothed_mean <- sample_mean[smoothed_index]
options(scipen = 10) # To stop the scientific method of counting
plot(smoothed_index, smoothed_mean, type = "l", col = "blue",
     xlab = "Number of Simulations (i)",
     ylab = "Running Mean of 95th Percentile Estimate",
     main = "Relatively smoothed Convergence Plot with 100 steps as the interval",
     ylim = c(1.5, 1.8))
abline(h = qnorm(0.95), col = "red", lty = 2)
```

For (b)

```{r}
n_sim <- 100000
simple_96th_vals <- numeric(n_sim)
quantile_96th_vals <- numeric(n_sim)
sample_mean <- numeric(n_sim)

# Simulation loop
for (i in 1:n_sim) {
  x <- rnorm(100)
  simple_96th_vals[i] <- simple_96th(x) # the  95th generation
  sample_mean[i] <- mean(simple_96th_vals[1:i])
}

# Summarize the results
cat("Mean of simple 96th estimate:", mean(simple_96th_vals), "\n")
```

```{r}
smoothed_index <- seq(100, 100000, by = 100)
smoothed_mean <- sample_mean[smoothed_index]
options(scipen = 10) # To stop the scientific method of counting
plot(smoothed_index, smoothed_mean, type = "l", col = "blue",
     xlab = "Number of Simulations (i)",
     ylab = "Running Mean of 95th Percentile Estimate",
     main = "Relatively smoothed Convergence Plot with 100 steps as the interval",
     ylim = c(1.5, 1.8))
abline(h = qnorm(0.95), col = "red", lty = 2)
```


```{r}
n_sim <- 100000
simple_9596th_vals <- numeric(n_sim)
quantile_9596th_vals <- numeric(n_sim)
sample_mean <- numeric(n_sim)

# Simulation loop
for (i in 1:n_sim) {
  x <- rnorm(100)
  simple_9596th_vals[i] <- simple_avg_9596(x) 
  sample_mean[i] <- mean(simple_9596th_vals[1:i])
}

# Summarize the results
cat("Mean of simple 95, 96th estimate:", mean(simple_9596th_vals), "\n")
```
```{r}
smoothed_index <- seq(100, 100000, by = 100)
smoothed_mean <- sample_mean[smoothed_index]
options(scipen = 10) # To stop the scientific method of counting
plot(smoothed_index, smoothed_mean, type = "l", col = "blue",
     xlab = "Number of Simulations (i)",
     ylab = "Running Mean of 95th Percentile Estimate",
     main = "Relatively smoothed Convergence Plot with 100 steps as the interval",
     ylim = c(1.5, 1.8))
abline(h = qnorm(0.95), col = "red", lty = 2)
```

For quantile()

```{r}
n_sim <- 100000
quantile_95th_vals <- numeric(n_sim)
sample_mean <- numeric(n_sim)

# Summarize the results

for (i in 1:n_sim) {
  x <- rnorm(100)
  quantile_95th_vals[i] <- quantile(x,0.95) 
  sample_mean[i] <- mean(quantile_95th_vals[1:i])
}

# Summarize the results
cat("Mean of quantile() estimate:", mean(quantile_95th_vals), "\n")
```


```{r}
smoothed_index <- seq(100, 100000, by = 100)
smoothed_mean <- sample_mean[smoothed_index]
options(scipen = 10) # To stop the scientific method of counting
plot(smoothed_index, smoothed_mean, type = "l", col = "blue",
     xlab = "Number of Simulations (i)",
     ylab = "Running Mean of 95th Percentile Estimate",
     main = "Relatively smoothed Convergence Plot with 100 steps as the interval",
     ylim = c(1.5, 1.8))
abline(h = qnorm(0.95), col = "red", lty = 2)
```



### (e)

The 95th estimate: 1.597052 
The 96th estimate: 1.692229 
The 95+96 /2 quantile() estimate: 1.639586 
The quantile() estimate : 1.596467 

```{r}
qnorm(0.95)
```

It seems like the 95+96/2 quantile estimate is the best method to estimate the 95th quantile. It better captures the interpolation that the true quantile lies between the 95th and 96th values, so it is a better estimator compared with the other three.  



## 2. Simulation of pi  

```{r}
pi2 <- function(n) {
  sum_val <- sum(1 / (1:n)^2)
  return(sqrt(6 * sum_val))
}
```


```{r}
for (j in 0:6) {
  n <- 10^j
  approximation <- pi2(n)
  cat("j =", j, ", n =", n, ", pi2(n) =", approximation, "\n")
}
```


```{r}
pi3 <- function(n) {
  x <- runif(n, -1, 1)
  y <- runif(n, -1, 1)
  inside <- (x^2 + y^2 <= 1)
  estimate <- 4 * mean(inside)
  return(estimate)
}
```


```{r}
set.seed(123)  # For reproducibility
for (j in 0:6) {
  n <- 10^j
  est <- pi3(n)
  cat("j =", j, ", n =", n, ", pi3(n) =", est, "\n")
}
```

The triangle means the points outside the circle and the circle represents the points that are inside the circle.

```{r}
n <- 10^4
x <- runif(n, -1, 1)
y <- runif(n, -1, 1)
inside <- (x^2 + y^2 <= 1)

plot(x[!inside], y[!inside], col = "blue", pch = 2, cex = 0.5,
     xlab = "x", ylab = "y", main = "Monte Carlo Points (j = 6)")
points(x[inside], y[inside], col = "violet", pch = 1, cex = 0.5)
legend("topright", legend = c("Outside circle", "Inside circle"),col = c("blue", "violet"), pch = c(3, 1), bty = "n")
```

```{r}
n <- 10^5
x <- runif(n, -1, 1)
y <- runif(n, -1, 1)
inside <- (x^2 + y^2 <= 1)

plot(x[!inside], y[!inside], col = "blue", pch = 2, cex = 0.5,
     xlab = "x", ylab = "y", main = "Monte Carlo Points (j = 6)")
points(x[inside], y[inside], col = "violet", pch = 1, cex = 0.5)
legend("topright", legend = c("Outside circle", "Inside circle"),col = c("blue", "violet"), pch = c(3, 1), bty = "n")
```


